Maximising the guaranteed feasible set for uncertain MPC schemes with chance constraints

### 2016-11-06 draft.tex
edits 3

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\newcommand{\Obs}{\mathcal{O}}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathcal C}
\providecommand{\E}{\mathcal E}
\providecommand{\F}{\mathscr F}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\Z}{\mathcal Z}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

\title{Maximising the guaranteed feasible set for uncertain MPC schemes with probabilistic constraints}%\thanksref{footnoteinfo}} 
% Title, preferably not more than 10 words.

@@

\begin{abstract}                          % Abstract of not more than 250 words.
This paper proposes a method of approximating positively invariant sets and $n$-step controllable sets of uncertain linear systems that are subject to chance constraints. The computed sets are robustly invariant and are guaranteed to satisfy the probabilistic constraints. In contrast, existing methods based on random sampling are only able to satisfy such constraints with a certain level of confidence. The proposed approach uses explicitly parametrised auxiliary disturbance sets, which are optimised so as to maximise the relevant positively invariant or $n$-step controllable set. The results are illustrated by numerical examples.
\end{abstract}

@@

\section{Introduction}
%
%
Positively invariant sets for uncertain linear systems have been researched for several decades, see e.g.~\cite{Kolmanovsky:1995,Kolmanovsky:1998,blanchini:2007}.
%
The properties of such sets and the algorithms to compute them are well known and documented in the literature.
%
As well as being objects of interest in their own right,
%In addition to pure analysis purposes, 
these sets are often used as terminal constraint sets model predictive control in order to guarantee the feasibility of the control scheme at all times \citep[see e.g.][]{Mayne2014}.
%

@@

Positively invariant sets based on samples of uncertain model parameters are often computationally expensive to determine and guarantees can only be given in a stochastic sense, namely with a certain confidence level.
%
The method presented in~\cite{Zhang2015} determines a set in which the probabilistic requirements are met with a given confidence and uses that set for subsequent analysis.
%
However, for simple probability distributions, this approach does not require sampling; instead we can explicitly parametrise an auxiliary disturbance set in which all probabilistic conditions are met.
%
Using such a parametrised set, we can perform an optimisation
%the size of the positively invariant set 
%for the particular set 
in order to compute the largest positively invariant set while guaranteeing satisfaction of the probabilistic constraints, hence approximating the maximal positively invariant set for a chance-constrained problem.
%
Unlike sample based approaches, this method ensures probabilistic constraint satisfaction with certainty (rather than with a confidence level less than $1$, which is the case for sample-based methods).
%
A similar principle of using parametrised sets of a specified probability measure is used in $p$-efficiency theory~\citep{Dentcheva2009}, where the distribution is not required to be simple but the structure of the chance constraints is restrictive.

In order to formulate a robust model predictive control law for a chance constrained stochastic system while introducing a minimal amount of conservatism, we consider a closed-loop min-max formulation~\citep{Lee:1997}. This requires the determination of additional constraints.
%
The $n$-step controllable set for a given horizon $n$ and target set \citep[see e.g.][]{bertsekas71} has the property that, for every state in the $n$-step controllable set, there exists an admissible input sequence such that the system state enters the target set after $n$ steps, for any admissible sequence of uncertainty realisations.
%
Using a parametrised uncertainty set with specified probabilistic properties, chance constraints can be replaced by computationally convenient set operations that guarantee the satisfaction of the original chance constraints.
%
By maximising the size of the $n$-step controllable set we indirectly approximate the $n$-step controllable set of the probabilistically constrained system.

%The remainder of this paper elaborates the method in the following way:
%
This paper is organised  as follows:
In section~\ref{sec:setup} we present the considered problem formulation and derive, in section~\ref{ssec:approximating:MRPI}, the conditions for the positively invariant set under chance constraints. The corresponding conditions for the $n$-step controllable set are derived in section~\ref{ssec:approximating:n:step}.
%
The parametrisation of a parallelotope is discussed in~\ref{sec:volume:of:hypercube} and the general problem formulations introduced in section~\ref{sec:setup} are made precise for this parametrisation.
%
Two numerical examples are presented in section~\ref{sec:examples} to illustrate the proposed scheme in two dimensions and in section~\ref{sec:conclusion} we conclude the presentation and sketch possible extensions to the presented scheme.

\textit{Notation.} Given sets $A,B\subset\RR^d$, $A\oplus B = \{x:x=a+b,\, a\in A,\, b\in B\}$ denotes the Minkowski sum and $A\ominus B = \{x: x+b \in A \ \forall b\in B\}$ the Pontryagin difference. We denote the image and pre-image of $C\subset\RR^d$ under the linear map $A:\RR^d\rightarrow\RR^d$ respectively as $AC = \{y\in\RR^d:Ax=y,\, x\in C\}$ and $A^{-1}C = \{y\in\RR^d:Ay\in C\}$. The $d$-dimensional volume of $C\subset\RR^d$ is denoted by $\vol(C)$. 
%
We define $v$ as a random parameter on a probability space $(\V,\F,\PP)$ and denote the probability of an event $E(v)$ as $\PP\{v\in\V : E(v)\}$.
%; throughout this paper the probability density of $v$ is assumed to be uniform.
%
The set $\B_P(r) = \{x\in\RR^d:x^TPx\leq r^2\}$ denotes the $P$-ball of radius $r$ centred at the origin. For a matrix $\Gamma\in\RR^{m\times n}$ the notation $\Gamma x\leq \bfa{1}$ is to be understood element-wise, where $\bfa{1}$ is a vector of ones in $\RR^m$.

@@

\subsection{Approximating the guaranteed invariant set}\label{ssec:approximating:MRPI}
%
First we discuss how to extend existing methods to calculate the maximal robust positively invariant (MRPI) set (see e.g.~\cite{blanchini:2007}) in such a way that all states in the MRPI set produce only feasible successor states for a given feedback controller $u = Kx$.
%
Thus we seek to determine a set of initial conditions for which all possible trajectories of (\ref{eq:system}) subject to (\ref{eq:disturbance}) satisfy the robust constraints~(\ref{eq:state_input:constraints}) and probabilistic constraint (\ref{eq:probabilistic:constraint}). Denoting the largest such set as $\X^\infty$ we have
\begin{multline*}
\X^\infty = \Bigl\{x\in\X: \\
\begin{aligned}
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i}\\
&x_k\in\X\\
&Kx_k\in\U\\
&\PP\{v\in\V : (C+DK)x_k+v\in\Y\}\geq p 
\end{aligned}\\
\forall w_k\in\W,\, \forall k\geq 0 \Bigr\} .
\end{multline*}
%
% \begin{multline}\label{eq:alternative:MRPI:set}
% 	X^\infty = \Bigl\{x\in\X: \\
% \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,\ \forall v\in \V^\prime,\ \forall k\geq 0 \Bigr\}
% \end{aligned} 
% \end{multline}
%
% It is obvious that $X^\infty$ contained in the invariant set of the original problem~$\X^\infty$:
For a given set $\V^\prime$, let $X^\infty$ denote the set
%
\begin{multline}\label{eq:alternative:MRPI:set}
	X^\infty = \Bigl\{x\in\X: \\
\begin{aligned} 
&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i} \\
&x_k\in\X \\ 
&Kx_k\in\U \\
&(C+DK)x_k+v_k\in\Y
\end{aligned}\\
\forall w_k\in\W,\ \forall v_k\in \V^\prime,\ \forall k\geq 0 \Bigr\} .
\end{multline}
% \begin{equation}
% 	\X^\infty = \left\{x\in\X: \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}
% \end{equation}
%
Then from Proposition~\ref{prop1}, the inclusion $X^\infty\subseteq\X^\infty$ necessarily holds for any $\V^\prime\subseteq\V$ with probability measure $\PP\{\V^\prime\} \geq p$. Furthermore, by maximising the volume of $X^\infty$, we can improve the approximation of $\X^\infty$.

@@

An optimisation program to approximate the positively invariant set $\X^\infty$ can therefore be formulated as
%
\begin{subequations}\label{seq:optimisation:MRPI:abstract}
	\begin{equation}
		\max_{\V^\prime} \vol(X^\infty)
	\end{equation}
%
	subject to
%
\begin{equation}
	\V^\prime\subseteq\V
\end{equation}
%
\begin{equation}
	\vol(\V^\prime)\geq p\cdot\vol(\V)
\end{equation}
\end{subequations}
%
In section~\ref{sec:volume:of:hypercube} we discuss how this optimisation can be solved using parallelotopes for~$\V^\prime$.

\subsection{Approximating the $n$-step controllable set}\label{ssec:approximating:n:step}
%
We now turn our attention to the $n$-step controllable set~$\C_n(\T)$ for a given target set~$\T$.
%
That is the set of initial conditions $x_0\in\C_n$ such that a sequence of control inputs exists that drives the state in $n$ steps to the target set, $x_n\in\T$, for all admissible disturbance sequences 
$w\in\W$, while guaranteeing that the robust state and input constraints~(\ref{eq:state_input:constraints}) and the probabilistic constraint~(\ref{eq:probabilistic:constraint}) are satisfied.
%
This set is given by
%
\begin{multline*}
\C_n(\T) = \Bigl\{x_0\in\X: \exists \bigl\{u_0(\cdot),\ldots,u_{n-1}(\cdot)\bigr\}\\
\begin{aligned}
&x_k = A^k x_0 + \sum_{i=1}^{k} A^{k-i}B u_{k-i}(x_{k-i}) + \sum_{i=1}^{k} A^{k-i} w_{k-i}\\
&x_n\in\T\\
&x_k\in\X\\
&u_k(x_k)\in\U\\
&\PP\bigl\{v\in\V : Cx_k + Du_k(x_k) +v \in\Y\bigr\}\geq p
\end{aligned}\\
\forall w_k\in\W,\, \forall k\in\{ 0, \ldots , n\} \Bigr\}	
\end{multline*}
% \begin{equation}
% 	\C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}	
% \end{equation}
%
Again, it is easy to verify that the set
%
\begin{multline*}
C_n(\T) = \Bigl\{ x\in\X: \exists \bigl\{u_0(\cdot),\ldots,u_{n-1}(\cdot)\bigr\}\\
\begin{aligned}
&x_k = A^k x_0 + \sum_{i=1}^{k} A^{k-i}B u_{k-i}(x_{k-i}) + \sum_{i=1}^{k} A^{k-i} w_{k-i}\\
&x_N\in\T\\
&x_k\in\X\\
&u_k(x_k)\in\U \\
&Cx_k+Du_k(x_k) + v_k \in\Y
\end{aligned}\\
\forall w_j\in\W,\, \forall v_k\in \V^\prime,\, \forall k\in\{0,\dots,n\} \Bigr\}	
\end{multline*}
% \begin{equation}
% 	C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,v\in \V^\prime,k\in\{0,\dots,N\}\end{aligned} \right\}	
% \end{equation}
%
is necessarily contained in $\C_n(\T)$ for any $\V^\prime\subseteq\V$ with measure $\PP\{\V^\prime\} \geq p$.
%
As in the case of the positively invariant set, fixing $\V^\prime$ to be a polytopic set leads to a computationally tractable optimisation for the maximisation of~$C_n$. This suggests an algorithm for constructing $C_n$ as a polytopic approximation of~$\C_n$ by optimising~$\V^\prime$ so as to maximise the volume of~$C_n$.
% indirectly improves the approximation of $\C_n$.
%
It is known that 
%$C_n(\T) = C^n_1(\T) \doteq C_1(C_1(\cdots C_1(\T)\cdots))$ 
$C_n(\T) = C^n_1(\T) \doteq C_1 \circ C_1 \circ \cdots \circ C_1(\T)$ 
\citep[see e.g.][]{blanchini:2007}, which allows the use of simple computational geometry algorithms.

In particular the one-step controllable set $C_1(\T)$ for 
\[
\T = \Bigl\{x\in\RR^d : J_i x\leq1,\ i\leq M_\T\Bigr\}
\]
is given by the projection of 
%
\begin{multline*}
L_1(\T) = \Bigl\{(x,u)\in (\X,\U): \\
\begin{aligned}
& Cx+Du + v\in\Y \\
& Ax+Bu+w\in\T
\end{aligned}\\
\forall v\in\V^\prime,\, \forall w\in\W\Bigr\}
%\end{aligned}
\end{multline*}
% \begin{equation}
% 	L_1(\T) = \left\{(x,u):\begin{aligned}
% 	x\in\X, u\in\U, Cx+Du + v\in\Y\;\forall v\in\V^\prime,\\
% 	Ax+Bu+w\in\T\;\forall w\in\W
% 	\end{aligned}\right\}
% \end{equation}
%
onto $\RR^d$.
%
As all involved sets are polytopic, $L_1(\T)$ is polytopic and its exact projection can be determined computationally.

The optimisation approximating the $n$-step controllable set $\C_n(\T)$ is then given as
%
\begin{subequations}\label{seq:optimisation:n:step:set:abstract}
	\begin{equation}
		\max_{\V^\prime} \vol(C_n(\T))
	\end{equation}
	subject to
	\begin{equation}
	\V^\prime\subseteq\V
\end{equation}
%
\begin{equation}
	\vol(\V^\prime)\geq p\cdot\vol(\V)
\end{equation}
\end{subequations}

\begin{rem}
The goal of reformulating a chance constrained model predictive control setup using the proposed method would be to minimise the introduced conservatism.
%
To achieve the best possible performance while guaranteeing constraint satisfaction a closed loop min-max model predictive control problem should be formulated, see e.g.~\cite{Lee:1997}.
%
This means that at each stage the decision variables are constrained to be such that the successor state is guaranteed to lie on a feasible trajectory to the target set.
%
Therefore to reformulate a chance constrained model predictive control problem as a deterministically constrained problem with guaranteed probabilistic feasibility and guaranteed recursive feasibility, the sets $C_n(\T)$ can be introduced as stage constraints, and the target set $\T$ can be chosen as the robust positively invariant set $X^\infty$.
\end{rem}

>>>>

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\newcommand{\Obs}{\mathcal{O}}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathscr C}
\providecommand{\Xinfty}{{\mathscr X}^\infty}
\providecommand{\E}{\mathcal E}
\providecommand{\F}{\mathscr F}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\Z}{\mathcal Z}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

\title{Maximising the guaranteed feasible set for uncertain MPC schemes with chance constraints}%\thanksref{footnoteinfo}} 
% Title, preferably not more than 10 words.

@@

\begin{abstract}                          % Abstract of not more than 250 words.
This paper proposes a method of approximating positively invariant sets and $n$-step controllable sets of uncertain linear systems that are subject to chance constraints. The computed sets are robustly invariant and are guaranteed to satisfy the probabilistic constraints of the control problem. In contrast, existing methods based on random sampling are only able to satisfy such constraints with a fixed level of confidence. The proposed approach uses explicitly parametrised auxiliary disturbance sets, which are optimised subject to a constraint on their probability measure so as to maximise the relevant positively invariant or $n$-step controllable set. The results are illustrated by numerical examples.
\end{abstract}

@@

\section{Introduction}
%
%
Positively invariant sets for uncertain linear systems have been researched for several decades, see e.g.~\cite{Kolmanovsky:1995,Kolmanovsky:1998,blanchini:2007}.
%
The properties of such sets and the algorithms to compute them are well known and documented in the literature.
%
As well as being objects of interest in their own right,
%In addition to pure analysis purposes, 
these sets are often used as terminal constraint sets in Model Predictive Control (MPC) in order to guarantee the feasibility of the control scheme at all times \citep[see e.g.][]{Mayne2014}.
%

@@

Positively invariant sets based on samples of uncertain model parameters are often computationally expensive to determine and guarantees can only be given in a stochastic sense, namely with a certain confidence level.
%
The method presented in~\cite{Zhang2015} determines a set in which the probabilistic requirements are met with a given confidence and uses that set for subsequent analysis.
%
However, for simple probability distributions such as the uniform distribution considered in this paper, this approach does not require sampling; instead we can explicitly parametrise an auxiliary disturbance set in which all probabilistic conditions are met.
%
Using such a parametrised set, we can perform an optimisation
%the size of the positively invariant set 
%for the particular set 
in order to compute the largest positively invariant set while guaranteeing satisfaction of the probabilistic constraints, hence approximating the maximal positively invariant set for a chance-constrained problem.
%
Unlike sample based approaches, this method ensures probabilistic constraint satisfaction with certainty (rather than with a confidence level less than $1$, which is the case for sample-based methods).
%
A similar principle of using parametrised sets of a specified probability measure is used in $p$-efficiency theory~\citep{Dentcheva2009}, where the distribution is not required to be simple but the structure of the chance constraints is restrictive.

In order to formulate a robust MPC law for a chance constrained stochastic system while introducing a minimal amount of conservatism, we consider a closed-loop min-max formulation~\citep{Lee:1997}. This requires the determination of additional constraints.
%
The $n$-step controllable set for a given horizon $n$ and target set \citep[see e.g.][]{bertsekas71} has the property that, for every state in the $n$-step controllable set, there exists an admissible input sequence such that the system state enters the target set after $n$ steps, for any admissible sequence of uncertainty realisations.
%
Using a parametrised uncertainty set with specified probabilistic properties, chance constraints can be replaced by computationally convenient set operations that guarantee the satisfaction of the original chance constraints.
%
By maximising the size of the $n$-step controllable set we indirectly approximate the $n$-step controllable set of the probabilistically constrained system.

%The remainder of this paper elaborates the method in the following way:
%
This paper is organised  as follows:
In section~\ref{sec:setup} we present the problem formulation and derive, in section~\ref{ssec:approximating:MRPI}, conditions that define the positively invariant set under chance constraints. The corresponding conditions for the $n$-step controllable set are derived in section~\ref{ssec:approximating:n:step} and these are used to define an MPC law in section~\ref{ssec:mpc}.
%
The parametrisation of a parallelotope is discussed in~\ref{sec:volume:of:hypercube} and the general problem formulations introduced in section~\ref{sec:setup} are made precise for this parametrisation.
%
Two numerical examples are presented in section~\ref{sec:examples} to illustrate the proposed scheme in two dimensions and in section~\ref{sec:conclusion} we conclude the presentation and sketch possible extensions to the presented scheme.

\textit{Notation.} Given sets $A,B\subset\RR^d$, $A\oplus B = \{x:x=a+b,\, a\in A,\, b\in B\}$ denotes the Minkowski sum and $A\ominus B = \{x: x+b \in A \ \forall b\in B\}$ the Pontryagin difference. We denote the image and pre-image of $C\subset\RR^d$ under the linear map $A:\RR^d\rightarrow\RR^d$ respectively as $AC = \{y\in\RR^d:Ax=y,\, x\in C\}$ and $A^{-1}C = \{y\in\RR^d:Ay\in C\}$. The $d$-dimensional volume of $C\subset\RR^d$ is denoted by $\vol(C)$. 
%
We define $v$ as a random parameter on a probability space $(\V,\F,\PP)$ and denote the probability of an event $E(v)$ as $\PP\{v\in\V : E(v)\}$.
%; throughout this paper the probability density of $v$ is assumed to be uniform.
%
For a positive definite matrix $P$ we write $\|x\|_P^2 = x^TPx$, and the set $\B_P(r) = \{x\in\RR^d:x^TPx\leq r^2\}$ denotes the $P$-ball of radius $r$ centred at the origin. The constraint $\Gamma x\leq \bfa{1}$, where $\Gamma\in\RR^{m\times n}$ and $\bfa{1}$ is a vector of ones in $\RR^m$, is understood to apply element-wise.

@@

\section{Approximating the guaranteed invariant set}\label{ssec:approximating:MRPI}
%
First we discuss how to extend existing methods to calculate the maximal robust positively invariant (MRPI) set (see e.g.~\cite{blanchini:2007}) in such a way that all states in the MRPI set produce only feasible successor states for a given feedback controller $u = Kx$.
%
Thus we seek to determine a set of initial conditions for which all possible trajectories of (\ref{eq:system}) subject to (\ref{eq:disturbance}) satisfy the robust constraints~(\ref{eq:state_input:constraints}) and probabilistic constraint (\ref{eq:probabilistic:constraint}). Denoting the largest such set as $\Xinfty$ we have
\begin{multline*}
\Xinfty = \Bigl\{x_0\in\X: \\
\begin{aligned}
% &x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i}\\
% &x_{k+1} = (A+BK)^{k+1} x + \sum_{i=1}^{k+1} (A+BK)^{k+1-i} w_{k+1-i}\\
&x_{k+1} = (A+BK)^{k+1} x_0 + \sum_{i=0}^{k} (A+BK)^{k-i} w_{i}\\
&x_{k+1}\in\X\\
&Kx_k\in\U\\
&\PP\{v\in\V : (C+DK)x_k+v\in\Y\}\geq p 
\end{aligned}\\
\forall w_k\in\W,\, \forall k\geq 0 \Bigr\} .
\end{multline*}
%
% \begin{multline}\label{eq:alternative:MRPI:set}
% 	X^\infty = \Bigl\{x\in\X: \\
% \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,\ \forall v\in \V^\prime,\ \forall k\geq 0 \Bigr\}
% \end{aligned} 
% \end{multline}
%
% It is obvious that $X^\infty$ contained in the invariant set of the original problem~$\Xinfty$:
For a given set $\V^\prime$, let $X^\infty$ denote the set
%
\begin{multline}\label{eq:alternative:MRPI:set}
	X^\infty = \Bigl\{x_0\in\X: \\
\begin{aligned} 
%&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i} w_{k-i} \\
&x_{k+1} = (A+BK)^{k+1} x_0 + \sum_{i=0}^{k} (A+BK)^{k-i} w_{i} \\
&x_{k+1}\in\X \\ 
&Kx_k\in\U \\
&(C+DK)x_k+v_k\in\Y
\end{aligned}\\
\forall w_k\in\W,\ \forall v_k\in \V^\prime,\ \forall k\geq 0 \Bigr\} .
\end{multline}
% \begin{equation}
% 	\Xinfty = \left\{x\in\X: \begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}
% \end{equation}
%
Then from Proposition~\ref{prop1}, the inclusion $X^\infty\subseteq\Xinfty$ necessarily holds for any $\V^\prime\subseteq\V$ with probability measure $\PP\{\V^\prime\} \geq p$. Furthermore, by maximising the volume of $X^\infty$, we can improve the approximation of $\Xinfty$.

@@

An optimisation problem whose solution approximates the maximal guaranteed positively invariant set $\Xinfty$ can therefore be formulated as
%
\begin{subequations}\label{seq:optimisation:MRPI:abstract}
\begin{align}
		& \max_{\V^\prime} \vol(X^\infty) \\
%\end{align}
%
\intertext{subject to}
%
%\begin{align}
	& \V^\prime\subseteq\V \\
% \end{equation}
% %
% \begin{equation}
	& \vol(\V^\prime)\geq p\cdot\vol(\V)
\end{align}
\end{subequations}
%
In section~\ref{sec:volume:of:hypercube} we discuss how this optimisation can be solved using a parallelotope parameterisation of~$\V^\prime$.

\section{Approximating the guaranteed $n$-step controllable set}\label{ssec:approximating:n:step}
%
We now turn our attention to the $n$-step controllable set~$\C_n(\T)$ for a given target set~$\T$.
%
This is defined as the set of initial conditions $x_0\in\X$ such that a sequence of control inputs exists that drives the state in $n$ steps to the target set, so that $x_n\in\T$ for all admissible disturbance sequences 
$w_k\in\W$, $k=0,\ldots,n-1$, while guaranteeing that the robust state and input constraints 
%$x_k\in\X$, $u_k\in\U$
(\ref{eq:state_input:constraints}) and the probabilistic constraint~(\ref{eq:probabilistic:constraint}) are satisfied.
%
Thus $\C_n(\T)$ can be expressed as follows.
%
\begin{multline*}
\C_n(\T) = \Bigl\{x_0\in\X: \exists \bigl\{u_0(\cdot),\ldots,u_{n-1}(\cdot)\bigr\}\\
\begin{aligned}
&x_{k+1} = A^{k+1} x_0 + \sum_{i=0}^{k} A^{k-i}B u_{i}(x_{i}) + \sum_{i=0}^{k} A^{k-i} w_{i}\\
&x_n\in\T\\
&x_{k+1}\in\X\\
&u_k(x_k)\in\U\\
&\PP\bigl\{v\in\V : Cx_k + Du_k(x_k) +v \in\Y\bigr\}\geq p
\end{aligned}\\
\forall w_k\in\W,\, \forall k\in\{ 0, \ldots , n-1\} \Bigr\}	
\end{multline*}
% \begin{equation}
% 	\C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&\PP\{(C+DK)x_k+v\in\Y\}\geq p\\
% 	&\forall w_j\in\W,k\geq 0\end{aligned} \right\}	
% \end{equation}
%
Using Proposition~\ref{prop1} it is easy to verify that the set $C_n(\T)$ defined by
%
\begin{multline*}
C_n(\T) = \Bigl\{ x_0\in\X: \exists \bigl\{u_0(\cdot),\ldots,u_{n-1}(\cdot)\bigr\}\\
\begin{aligned}
&x_{k+1} = A^{k+1} x_0 + \sum_{i=1}^{k} A^{k-i}B u_{i}(x_{i}) + \sum_{i=1}^{k} A^{k-i} w_{i}\\
&x_N\in\T\\
&x_{k+1}\in\X\\
&u_k(x_k)\in\U \\
&Cx_k+Du_k(x_k) + v_k \in\Y
\end{aligned}\\
\forall w_k\in\W,\, \forall v_k\in \V^\prime,\, \forall k\in\{0,\dots,n-1\} \Bigr\}	
\end{multline*}
% \begin{equation}
% 	C_n(\T) = \left\{x\in\X:\begin{aligned}&x_k = (A+BK)^k x + \sum_{i=1}^{k} (A+BK)^{k-i-1} w_{k-i}\\
% 	&x_N\in\T\\
% 	&x_k\in\X\\
% 	&(C+DK)x_k+v\in\Y\\
% 	&\forall w_j\in\W,v\in \V^\prime,k\in\{0,\dots,N\}\end{aligned} \right\}	
% \end{equation}
%
is necessarily contained in $\C_n(\T)$ for any $\V^\prime\subseteq\V$ with probability measure $\PP\{\V^\prime\} \geq p$.
%
As in the case of positively invariant sets, fixing $\V^\prime$ to be a polytopic set leads to a computationally tractable optimisation for the maximisation of~$C_n$. This suggests an algorithm for constructing $C_n$ as a polytopic approximation of~$\C_n$ by optimising~$\V^\prime$ so as to maximise the volume of~$C_n$.
% indirectly improves the approximation of $\C_n$.
%
It is known that 
%$C_n(\T) = C^n_1(\T) \doteq C_1(C_1(\cdots C_1(\T)\cdots))$ 
$C_n(\T) = C^n_1(\T) \doteq C_1 \circ C_1 \circ \cdots \circ C_1(\T)$ 
\citep[see e.g.][]{blanchini:2007}, which allows the use of simple computational geometry algorithms.

In particular, the one-step controllable set $C_1(\T)$ for the target set 
\[
\T = \Bigl\{x\in\RR^d : J_i x\leq1,\ i\leq M_\T\Bigr\}
\]
is given by the projection of
$L_1(\T)\subset \RR^{d+q_{\U}}$, 
%$L_1(\T)\subset \RR^d\times \RR^{q_{\U}}$, 
defined by
%
\begin{multline*}
L_1(\T) = \Bigl\{(x,u)\in (\X,\U): \\
\begin{aligned}
& Cx+Du + v\in\Y \\
& Ax+Bu+w\in\T
\end{aligned}\\
\forall v\in\V^\prime,\, \forall w\in\W\Bigr\} ,
%\end{aligned}
\end{multline*}
% \begin{equation}
% 	L_1(\T) = \left\{(x,u):\begin{aligned}
% 	x\in\X, u\in\U, Cx+Du + v\in\Y\;\forall v\in\V^\prime,\\
% 	Ax+Bu+w\in\T\;\forall w\in\W
% 	\end{aligned}\right\}
% \end{equation}
%
onto the $x$-subspace.
%
Thus
\[
C_1(\T) = \{ x : \exists u \ \text{such that} \ (x,u)\in L_1(\T)\}.
\]
As all involved sets are polytopic, $L_1(\T)$ is polytopic and its exact projection can be determined computationally.

A set $C_n(\T)$ approximating the $n$-step controllable set $\C_n(\T)$ is then given by the solution of the optimisation:
%
\begin{subequations}\label{seq:optimisation:n:step:set:abstract}
\begin{align}
	& \max_{\V^\prime} \vol(C_n(\T))
%\end{equation}
\intertext{subject to}
%\begin{align}
	& \V^\prime\subseteq\V  \\
% \end{equation}
% %
% \begin{equation}
	& \vol(\V^\prime)\geq p\cdot\vol(\V)
\end{align}
\end{subequations}
The property $C_n(\T)\subseteq \C_n(\T)$ then guarantees the existence of a control sequence such that the constraints~(\ref{eq:state_input:constraints}) and (\ref{eq:probabilistic:constraint}) are met and $x_n\in\T$ for all $x_0\in C_n(\T)$.

Note that an additional improvement to the $n$-step controllable set $C_n(\T)$ could be obtained by introducing $\V_1^\prime,\dots,\V_{n-1}^\prime$, to allow different sets $\V_i^\prime$ at each stage.
% introducing $(n-1)\times 2d$ additional decision variables.

\section{Model Predictive Control}\label{ssec:mpc}

The method proposed for computing $n$-step controllable sets in section~\ref{ssec:approximating:n:step} can be used to formulate a chance constrained Model Predictive Control law. In order to minimise conservatism while guaranteeing constraint satisfaction, a closed loop MPC scheme should be used \citep[see e.g.][]{Lee:1997}.
This involves computing (offline) the set $X^\infty$ using \mbox{(\ref{seq:optimisation:MRPI:abstract}a-c)}, and the 
sets $C_{i}(X^\infty) = C_1\bigl(C_{i-1}(X^\infty)\bigr)$ and $L_{i}(X^\infty) = L_1\bigl(C_{i-1}(X^\infty)\bigr)$ 
for $i=1,\ldots,n$ by repeated application of (\ref{seq:optimisation:n:step:set:abstract}a-c) with $C_0(X^\infty)= X^\infty$.
Online, at each time step $k=0,1,\ldots$, a min-max MPC law is defined by 
\[
u_k = u^{(n)}(x_k)
\]
where $u^{(n)}(x_k)$ is the solution of
%
\[%\begin{equation}
\begin{aligned}
\Bigl( u^{(n)} (x_k) , w^{(n)} (x_k,u) \Bigr) &= \arg \min_{u} \Bigl\{ \max_{w\in\W} J^{(n)}(x_k,u,w)  
\\
&\ \quad \text{subject to}\ (x_k, u) \in L_n(\T)
\Bigr\}
\end{aligned}
\]%\end{equation}
and $J^{(n)}$ is defined by the recursion, for $i=1,\ldots,n$:
\begin{align*}
&J^{(i)}(x,u,w) = \tfrac{1}{2}\bigl( \| x \|^2_Q + \| u \|^2_R -
\gamma^2 \| w \|^2\bigr) + J^{(i-1)}(x^+) 
\\
&J^{(i-1)}(x) = J^{(i-1)}\bigl(x,u^{(i-1)}(x),w^{(i-1)}(x,u^{(i-1)}(x))\bigr) \\
&x^+ = Ax + Bu + Dw
\end{align*}
with terminal cost $J^{(0)} = \frac{1}{2}\|x\|_{P_0}^2$.
%for some suitable terminal weighting matrix $P_0$.

An algorithm for solving the online optimization of this min-max robust MPC law is given in~\citet{Buerger2016}. 
%
Since the stage constraints are constructed using the controllable sets $C_{i}(X^\infty)$, the decision variables are constrained at each stage to be such that the successor state is guaranteed to lie on a feasible trajectory to the target set $X^\infty$. Therefore the MPC law solves a deterministically constrained problem with guaranteed probabilistic feasibility and guaranteed recursive feasibility.
